# Process PVZ data

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE, 
  warning = FALSE, 
  cache = FALSE
)
# Create directories for saving data
dir.create("data/ea/", FALSE, TRUE)
dir.create("data-raw/ea/", FALSE, TRUE)
```

Here, we process the Plants vs. Zombies: Battle for Neighborville (PVZ) survey and telemetry data files (i.e., gaming data that was captured by EA's servers).

We used the following R packages. Pacman is a package manager that install packages if you don't have them installed. We always show the package versions at the bottom of each section.

```{r packages}
library(pacman)
p_load(
  readxl,
  knitr,
  here,
  lubridate,
  scales,
  janitor,
  anytime,
  tidyverse
)
```

## Raw data

The raw data files, including the raw survey data, are in a zip compressed file on OSF: <https://osf.io/cjd6z/>. (A user ID column was manually removed from the survey data before uploading.) Start by downloading that file and unpacking it to the target directory:

```{r}
file_dest <- here("data-raw/ea/ea.zip")
# Download & extract file only if you haven't yet
if (!file.exists(file_dest)) {
  download.file("https://osf.io/4gp3r/download", file_dest)
}
if (!file.exists(here("data-raw/ea/Oxford PvZ - Pilot Wave + Wave 1 Download.xlsx"))) {
  unzip(file_dest, exdir = here("data-raw/ea/"))
}
```

EA sent out the survey in two waves: one in August of 2020, the second in September 2020. For each wave, there are 9 telemetry files, each one recording a different aspect of play events in Plants vs. Zombies: Battle for Neighborville and one file for the survey results.

These telemetry files contain info on PVZ play events from a long period of time, including for players who did not participate in the survey. We limit our analyses to the individuals who participated in the survey, and to the 2 weeks preceding the survey. However, the raw data files might be of interest for additional analyses.

## Process raw files

Because there were two of each telemetry file (waves), we create a function to read both waves' data into a single table:

```{r}
read_both_waves <- function(stub = "OII_PVZ_Authentications", ...) {
  # Filenames of wave 1 and 2 authentications files
  list.files(
    here("data-raw/ea"), pattern = stub, full.names = TRUE
  ) %>% 
    # Put wave 1 first
    rev() %>% 
    # Read into a list of data frames
    map(read_csv, ...) %>% 
    # Bind into one data frame with wave ID
    bind_rows(.id = "wave") %>%
    # Take only participants who filled out survey
    filter(player_id %in% survey$player_id)
}
```

### Survey

Because we only keep telemetry data of players who answered the survey, we'll load the survey first.
```{r}
survey <- read_excel(
  here("data-raw/ea/Oxford PvZ - Pilot Wave + Wave 1 Download.xlsx")
)

survey <- survey %>% 
  mutate(
    # Ensure date data type
    date = mdy_hm(date),
    # Need another name to use in filtering telemetry
    start_date = date
  ) %>% 
  # Other files use player_id
  rename(player_id = UID)
```

#### Clean and transform

Next, we give some sensible variable names and assign proper variable types for the survey.
```{r give-proper-variable-names}
# Rename variables
survey <- 
  survey %>% 
  select(
    
    # meta-info
    player_id,
    start_date, date,
    survey_duration = qtime,
    of_age = Q2,
    consent = Q3,
    
    # demographics
    played = Q1,
    country = hCountry,
    survey_language = PRELANG,
    age = D1,
    gender = D2,
    gender_other = D2r3oe,
    experience = D3,
    
    # well-being
    spane_ = starts_with("A1"), # general well-being in the two weeks
    straightliner_spane = hA1Straight,
    
    # motivations
    played_with_others = NSat1, # filter question that determines whether relatedness items were shown
    autonomy_ = NSat2r1:NSat2r3,
    competence_ = NSat2r4:NSat2r6,
    relatedness_ = NSat2r7:NSat2r9,
    enjoyment_ = NSat2r10:NSat2r13,
    extrinsic_ = NSat2r14:NSat2r17,
    straightliner_motivations = hNSat2Straight,
    
    # self-reported play
    active_play_hours = GT1c1,
    active_play_minutes = GT1c2,
    within_estimate = GT2,
    between_estimate = GT3,
    spane_game_ = starts_with("GT4")
  )

# Assign proper variable types and factor levels
survey <- survey %>%   
  mutate(
    across(
      c(
        player_id,
        of_age:survey_language,
        gender,
        played_with_others
      ),
      as.factor
    ),
    
    across(
      c(of_age:played, played_with_others),
      ~ fct_recode(
        .x,
        "Yes" = "1",
        "No" = "2"
      )
    ),
    
    gender = fct_recode(
      gender,
      "Male" = "1",
      "Female" = "2",
      "Other" = "3",
      "Prefer not to say" = "4"
    ),
    
    country = fct_recode(
      country,
      "US" = "1",
      "UK" = "2",
      "Canada" = "5"
    ),
    
    survey_language = fct_recode(
      survey_language,
      "English" = "1",
      "French" = "2"
    )
  )

# Reverse scored items
survey <- survey %>% 
  mutate(
    across(
      c(
        starts_with("spane"), # 1 was never
        between_estimate, #  those estimates also had 1: much more than typically
        within_estimate,
        relatedness_3,
        enjoyment_2,
        enjoyment_3
      ),
      ~ 8 - .x
    )
  )

# Indicators for straightliners
survey <- survey %>% 
  mutate(across(contains("straightliner"), ~.x==1))
```

Then we create mean indices for the scales.
SPANE has positive affect, negative affect, and an affect balance score (subtract negative from positive).
```{r create-scales}
# General SPANE
survey <- 
  survey %>% 
  mutate(
    spane_positive = rowMeans(
      select(
        .,
        spane_1,
        spane_3,
        spane_5,
        spane_7,
        spane_10,
        spane_12
      ),
      na.rm = TRUE
    ),
    spane_negative = rowMeans(
      select(
        .,
        spane_2,
        spane_4,
        spane_6,
        spane_8,
        spane_9,
        spane_11
      ),
      na.rm = TRUE
    ),
    spane_balance = spane_positive - spane_negative
  )

# Motivations
survey <- survey %>% 
  mutate(
    autonomy = rowMeans(
      select(., starts_with("autonomy")), na.rm = TRUE
    ),
    competence = rowMeans(
      select(., starts_with("competence")), na.rm = TRUE
    ),
    relatedness = rowMeans(
      select(., starts_with("relatedness")), na.rm = TRUE
    ),
    enjoyment = rowMeans(
      select(., starts_with("enjoyment")), na.rm = TRUE
    ),
    extrinsic = rowMeans(
      select(., starts_with("extrinsic")), na.rm = TRUE
    )
  )

# SPANE because of playing game
survey <- survey %>% 
  mutate(
    spane_game_positive = rowMeans(
      select(
        .,
        spane_game_1,
        spane_game_3,
        spane_game_5,
        spane_game_7,
        spane_game_10,
        spane_game_12
      ),
      na.rm = TRUE
    ),
    spane_game_negative = rowMeans(
      select(
        .,
        spane_game_2,
        spane_game_4,
        spane_game_6,
        spane_game_8,
        spane_game_9,
        spane_game_11
      ),
      na.rm = TRUE
    ),
    spane_game_balance = spane_game_positive - spane_game_negative
  )

# Hours of estimated play
survey <- survey %>% 
  mutate(
    active_play = active_play_hours + (active_play_minutes / 60)
  )
```

### Game time

Then we get to the telemetry files.

The game time data are the most complicated. We work on it first because the times are needed for some of the other telemetry files.
Note: The raw excel files on the OSF (or downloaded as zip files above) have a detailed data map as the second tab of the file. 
These serve as a codebook.
The data show us when a player started a session and how long that session lasted, plus whether they played the game in multiplayer or single player.
The variables:

* `player_id` = Unique identifier for a player.
* `platform` = One of the major platforms, eg. PC, Playstation, Xbox.
* `game_start_time` = Session start date time.
* `game_end_time` = Session end date time.
* `game_mode` = The game mode, primarily multiplayer or single player.
* `game_type` = Notes if the game is offline/online and if it is splitscreen
* `game_level` = The level's name where the session took place.
* `game_session` = Unique identifier for a game session.

```{r cache = TRUE}
game_time <- read_both_waves("OII_PVZ_Game_Time")
glimpse(game_time)
```

A game session starts when a player starts playing.
That means `game_session` records when a match starts, which includes going to the Hub World (i.e., a world inside the game where players can decide what to do next).
Game sessions can be shared between players when they're playing multiplayer, but should be unique to a `player_id` when in single player.
However, it's not easy to say when a session ends for technical reasons.
Ideally, we'd have one game session per row.
Because it's hard to determine when a session ends, there are instances where a given player has multiple rows with the same `game_session`.

```{r}
counts <- game_time %>% 
  count(player_id, game_session, sort = T)
```

For example, the player below has `r counts %>% slice(1) %>% pull(n)` rows with identical `player_id` and `game_session`.
Some of those rows have identical start and end times for the session, but the majority are in "chunks" with the same start times, but slightly different end times.
This pattern shows the problems with determining the end of a session.
See the figure below to see how the rows overlap in their game durations.
```{r}
# example of one player_idXgame_session with multiple entries
tmp <- counts %>% 
  slice(1) %>% 
  left_join(game_time) %>%
  arrange(game_start_time, game_end_time)

tmp %>%   
  rowid_to_column() %>% 
  ggplot(aes(game_start_time, rowid, col = game_session)) +
  geom_segment(aes(xend = game_end_time, yend = rowid)) +
  facet_wrap("player_id", labeller = label_both) +
  theme(legend.position = "bottom")
```

For those instances where a player has multiple (overlapping) game sessions, we'll simply use the earliest start time and latest end time of that "super-session" to collapse all overlapping rows into one.
For that, we look for each combination of `player_id` and `game_session` and replace `game_start_time` with the earliest time within that super-session; likewise, we replace `game_end_time` with the latest time within that super-session.
If a player only has one row, they'll maintain that one row this way.
If a player has multiple rows, they're collapsed into one row that encompasses the game durations of all rows.
See the figure below which now only has one duration for that player, which goes from the minimum to the maximum time of the previous figure.
```{r}
# Key point: Replace start and end times with first and last start/end time
#  then take unique rows. This results in 1 row per session.
tmp <- tmp %>% 
  group_by(player_id, game_session) %>% 
  mutate(
    game_start_time = min(game_start_time),
    game_end_time = max(game_end_time)
  ) %>% 
  slice(1) # distinct() didn't work when there were different levels within the same session
last_plot() %+% mutate(tmp, rowid = 1:n())
```

Next, we apply this procedure to all players.
However, some start times are missing, which won't allow us to compute duration for that session.
Whenever a `game_start_time` is missing, we'll set it to the end time of that session.
That creates game sessions with a duration of 0, which we'll set to `NA` again later.
(Note: `game_end_time` doesn't have missing values, which is why we can use `coalesce`)
```{r}
game_time <- game_time %>% 
  mutate(
    game_start_time = coalesce(
      game_start_time, game_end_time
    )
  )

# now collapse the rows
game_time <- game_time %>% 
  group_by(player_id, game_session) %>% 
  mutate(
    game_start_time = min(game_start_time),
    game_end_time = max(game_end_time),
  ) %>% 
  # Take the first of each (identical) person-session combo
  slice(1) %>% 
  ungroup()
```

Then we save the file.
```{r}
write_rds(game_time, here("data/ea/game_time.rds"))
```

### Authentications

The authentifications data set logs events when a player (`player_id`) logs into the game.
The variables are:

* `player_id` = Unique identifier for a player.
* `platform` = One of the major platforms, eg. PC, Playstation, Xbox.
* `date` = Date of the login.
* `country` = Country the login came from.

```{r cache = TRUE}
authentications <- read_both_waves("OII_PVZ_Authentications")
glimpse(authentications)
```

First, we only keep authentications of players who also participated in the survey.
Then we count how many times players logged in to PvZ.
In the survey, players reflected on their well-being and gaming in the past two weeks, so we'll only count those game events that occurred within 14 days prior to the time the participant took the survey.
The `date` variable in `authentications` doesn't have time, so we'll make it easy and also reduce `start_time` in the survey to a date (without time of day).
```{r}
authentications <- authentications %>% 
  # add the start date by player ID
  left_join(select(survey, player_id, start_date)) %>% 
  # only keep authentication events that occured within 14 days before the survey was taken
  filter(date >= (as_date(start_date) - days(14)) ) %>% 
  count(player_id, name = "authentications")
```

### Characters

The characters data provides performance stats for a player while they play as a specific character in the game.
The variables are:

* `player_id`
* `platform` = One of the major platforms, eg. PC, Playstation, Xbox.
* `game_session` = Unique identifier for a game session.
* `game_mode` = The game mode, primarily multiplayer or single player.
* `game_type` = Notes if the game is offline/online and if it is splitscreen
* `game_level` = The level's name where the session took place.
* `character_class` = The character key noting the character used.
* `total_kill_count` = Total kills earned by the player.
* `total_death_count` = Total deaths suffered by the player.
* `score` = Total score earned by the player.
* `damage_dealt` = Total damage dealt by the player.
* `critical_hit_count` = Number of critical hits earned by the player.
* `shots_fired` = Number of shots fired by the player.
* `shots_hit` = Number of shots by the player which successfully hit.

```{r cache = TRUE}
characters <- read_both_waves("OII_PVZ_Character")
glimpse(characters)
```

We do the same for the performance stats in the `characters` data set.
The data set doesn't have a start time for the game sessions, so we'll add those start times by matching the `player_id` and `game_session` with those of the `game_time` data set, which has start times for the game sessions, thereby adding the `game_start_time` for a game session.
Then we include only those game sessions that happened within 14 days of a participant taking the survey.
After that, we sum up all player stats for that period.
```{r}
characters <- characters %>%
  left_join(
    game_time %>% 
      select(player_id, game_session, game_start_time),
    by = c("player_id", "game_session")
  ) %>%
  left_join(select(survey, player_id, start_date)) %>% 
  filter(game_start_time >= (start_date - days(14))) %>% 
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>% 
  ungroup()
```

### Friends

The friends data show when two players became friends and at what time (in PvZ, not on the platform?).
The variables:

* `player_id` = Unique identifier for a player.
* `friend_player_id` = Unique identifier for a player, who is marked as a friend of the player_id.
* `date_time` = The date time when the friend invite was accepted.

```{r cache = TRUE}
friends <- read_both_waves("OII_PVZ_Friends")
write_rds(friends, here("data/ea/friends.rds"))
glimpse(friends)
```

Then we count the number of friend connections made.
`player_id` is the player; `friend_player_id` is the player who was sent a friend request from `player_id`.
That means some players will be in the `player_id` column when they send a friend request, but also in the `friend_player_id` column when they receive a friend request.
Therefore, we sum over sent and received friend connections in the 14 days prior to the survey.

```{r}
# filter out those that didn't happen within the past two weeks
friends <- friends %>% 
  left_join(select(survey, player_id, start_date)) %>% 
  filter(event_date_time >= (start_date - days(14)))

# get friend requests sent
a <- friends %>%
  group_by(player_id) %>% 
  summarise(requests_sent = length(unique(friend_player_id)))

# get friend requests received
b <- friends %>%
  group_by(friend_player_id) %>% 
  summarise(requests_received = length(unique(player_id)))

# add those new variables to the data
friends <- left_join(a, b, by = c("player_id" = "friend_player_id"))

# not a lot of people received friend requests (meaning they're NA), so before adding sent and received up, we set NAs to zero
friends <- friends %>%
  mutate(
    across(
      c(starts_with("requests")),
      ~ if_else(is.na(.x), 0L, .x)
    ),
    friends = requests_sent + requests_received
  ) %>% 
  select(-starts_with("requests"))

rm(a, b)
```


### Gestures

The gesture data shows when a player used a gesture (gesture/emote/command) that can be seen by other players.
The variables:

* `player_id` = Unique identifier for a player.
* `gesture_key` = Key name of the geature used.
* `gesture_type` = Whether the gesture was a gesture, command or emote.
* `game_session` = Unique identifier for a game session.
* `date_time` = The date time when the gesture was used.

```{r cache = TRUE}
gestures <- read_both_waves("OII_PVZ_Gesture")
write_rds(gestures, here("data/ea/gestures.rds"))
glimpse(gestures)
```

We first match only those sessions that were within the time frame, then count how many gestures a player used.

```{r}
gestures <- gestures %>% 
  left_join(select(survey, player_id, start_date)) %>% 
  filter(date_time >= (start_date - days(14))) %>% 
  count(player_id, name = "gestures")
```

### Leveling

This data set shows when a player went up a level with a character.
The variables:

* `player_id` = Unique identifier for a player.
* `platform` = One of the major platforms, eg. PC, Playstation, Xbox.
* `level` = The level earned for the character.
* `date_time` = The date time when the level was earned.
* `character_class` = The character key noting the character used.

```{r cache = TRUE}
leveling <- read_both_waves("OII_PVZ_Leveling")
write_rds(leveling, here("data/ea/leveling.rds"))
glimpse(leveling)
```

We only include those sessions that happened within the time frame, then we count how many times a player went up a level.

```{r}
leveling <- leveling %>%
  left_join(select(survey, player_id, start_date)) %>% 
  filter(date_time >= (start_date - days(14))) %>% 
  count(player_id, name = "levelings")
```

### Prestige

This data set show when a a player went up a prestige level with a character. 
Prestige levels are different from character levels.
The variables:

* `player_id` = Unique identifier for a player.
* `platform` = One of the major platforms, eg. PC, Playstation, Xbox.
* `prestige_level` = The prestige level for the character.
* `date_time` = The date time the prestige level was earned.
* `character_class` = The character key noting the character used.

```{r cache = TRUE}
prestige <- read_both_waves("OII_PVZ_Presti")
glimpse(prestige)
write_rds(prestige, here("data/ea/prestige.rds"))
```

```{r}
prestige <- prestige %>%
  left_join(select(survey, player_id, start_date)) %>% 
  filter(date_time >= (start_date - days(14))) %>% 
  count(player_id, name = "prestige")
```

### Registrations

Registrations only happen once per player and show when a player registered with PvZ for the first time.
Won't be needed later, but code is here for completeness.
The variables:

* `player_id` = Unique identifier for a player.
* `platform` = One of the major platforms, eg. PC, Playstation, Xbox.
* `date` = Date of the registration.
* `country` = Country the login came from.

```{r eval = FALSE}
registrations <- read_both_waves("OII_PVZ_Registration")
```

### Experience Points

This data set shows how many XP a player earned at what time and in what session.
The variables:

* `player_id` = Unique identifier for a player.
* `platform` = One of the major platforms, eg. PC, Playstation, Xbox.
* `date_time` = The date time when the xp was earned.
* `character_class` = The character key noting the character used.
* `game_session` = Unique identifier for a game session.
* `xp_earned` = The xp amount earned.

```{r cache = TRUE}
# Gotta process dates because they are different across the two files
xp <- read_both_waves(
  "OII_PVZ_XPEarned", 
  col_types = cols(date_time = "c")
)

# All other timestamps, above, are automatically parsed to UTC so we do that here as well
xp <- xp %>% 
  mutate(date_time = anytime(date_time, "UTC"))
glimpse(xp)
write_rds(xp, here("data/ea/xp.rds"))
```

We sum up how many experience points a player got in total during the two weeks before the survey.

```{r}
xp <- xp %>% 
  left_join(select(survey, player_id, start_date)) %>% 
  filter(date_time >= (start_date - days(14))) %>% 
  group_by(player_id) %>% 
  summarise(xp = sum(xp_earned)) %>% 
  ungroup()
```

### Process game time

The raw game times were used in processing the other files, so we left processing it until here.

First, we check which game sessions happened within the 14 days a participant took a survey (the file is already filtered to only include participants who actually filled out the survey). 

Afterwards, we aggregate the remaining game sessions per participant to get to how many hours they played in the two weeks before the survey (i.e., only aggregate/summarize `game_session`s that happened within the past 14 days).

```{r}
game_time <- game_time %>% 
  left_join(select(survey, player_id, start_date)) %>% 
  filter(game_start_time >= (start_date - days(14))) %>% 
  select(player_id, game_start_time, game_end_time) %>% 
  mutate(duration = game_end_time-game_start_time) %>% 
  mutate(Hours = as.numeric(duration)/60/60) %>% 
  group_by(player_id) %>% 
  summarise(Hours = sum(Hours)) %>% 
  ungroup()
```

## Merge survey and telemetry

Next, we add the person-level aggregated telemetry variables to the survey data.
The resulting data set will not deal with session level data, but summarizes, for each participant, their engagement in the 2 week observation period.

First put all the aggregated telemetry into one data frame.

```{r}
telemetry <- reduce(
  list(authentications, characters, friends, game_time, gestures, leveling, prestige, xp),
  left_join
)
```

And then merge them to the survey.
```{r}
pvz <- survey %>% 
  left_join(telemetry, by = "player_id")
```

## Exclusions

First save a file with no exclusions.

```{r}
write_rds(pvz, here("data/ea/pvz.rds"))
```

### Straightliners

We take out all players who straightlined (gave the same response to every item) through SPANE and motivations scales. (If only SPANE items existed, then we didn't exclude.)

```{r}
pvz <- pvz %>% 
  mutate(
    straightliner = 
      straightliner_spane & straightliner_motivations
  ) 
pvz %>%   
  select(contains("straight")) %>% 
  group_by_all() %>% 
  count
pvz %>%   
  tabyl(straightliner) %>% 
  adorn_pct_formatting()
# filter() would also exclude NAs!
pvz <- filter(pvz, !straightliner | is.na(straightliner))
```

### Outliers

Potential outliers. We replace all values that are more than 6SD away from the variable's mean with NAs. As a consequence, individuals are excluded on an analysis-by-analysis case (so if has bad data relevant to that analysis or figure).

This is only done for a subset of variables (relavant to analyses; see below).

```{r}
pvz <- pvz %>% 
  # These variables will be affected
  pivot_longer(
    c(spane_1:xp, -played_with_others, -contains("straightliner"))
  ) %>% 
  group_by(name) %>% 
  mutate(z_value = as.numeric(scale(value))) 
```

These are the numbers of people taken out of each variable (only variables that were affected are shown):

```{r}
# This is what are taken out
pvz %>% 
  summarise(
    Extremes = sum(abs(z_value>=6), na.rm = TRUE),
    Extremes_p = percent(Extremes/n(), accuracy = .01)
  ) %>% 
  filter(Extremes > 0)
```

Code to do it:

```{r}
pvz <- pvz %>%
  mutate(value = ifelse(abs(z_value >= 6), NA, value)) %>% 
  select(-z_value) %>% 
  pivot_wider(names_from = "name", values_from = "value") %>% 
  ungroup()
```


## Save files

```{r}
write_rds(pvz, here("data/ea/pvz-excluded.rds"))
```

## Session info
```{r}
sessionInfo()
```

