# Process AC:NH data

```{r setup, include=FALSE}
library(pacman)
p_load(
  readxl,
  knitr,
  here,
  lubridate,
  tidyverse
)
# remotes::install_github("rubenarslan/formr")
p_load(formr)
opts_chunk$set(
  echo = TRUE, 
  message = FALSE, 
  warning = FALSE, 
  cache = TRUE
)
```

Here, we process the AC:NH survey and telemetry files.

Currently this script pulls survey data from formr.org, and saves to `data-raw/noa/`. If data already exists there, then it is loaded instead.

Telemetry data doesn't exist yet.

After processing, we place the processed data sets in `data/noa/` as R data objects.

## Process formr surveys

### Get raw data

Ensure output directories exists

```{r}
dir.create("data-raw/", FALSE)
dir.create("data-raw/noa/", FALSE)
dir.create("data/", FALSE)
dir.create("data/noa/", FALSE)
```

Read data from formr.org and save to `data-raw/` if it doesn't yet exist. For now you have to manually authenticate to formr if needed, sorry.

```{r}
if (!file.exists(here("data-raw/noa/formr.rds"))) {
  formr_connect()
  # These are the surveys
  surveys <- c("gaming_welcome", paste0("gaming_block", 2:5))
  # List of surveys' data frames
  ac <- map(surveys, ~formr_results(.) %>% as_tibble)
  saveRDS(ac, here("data-raw/noa/formr.rds"))
} else(ac <- readRDS(here("data-raw/noa/formr.rds")))
```

Take only responses after the survey went live (i.e. take out our test sessions). Later, to make sure, we need to retain only rows where `code` exists in telemetry.

```{r}
ac[[1]] <- ac[[1]] %>% filter(created >= ymd("2020-10-27"))
```

Then, retain only rows where code exists, and consent & age check out

```{r}
# Keep only surveys whose code exists in gaming_welcome$code
ac[[1]] <- filter(
  ac[[1]], !is.na(code), of_age==1, consent_data==1, consent==1
)
```

Now apply those exclusions to all five survey data frames

```{r}
ac <- map(ac, ~filter(., code %in% ac[[1]][["code"]]))
```

Take out all time variables except first created and last ended (when survey started and ended)

```{r}
ac[[1]] <- select(ac[[1]], -c(modified:expired))
ac[2:4] <- map(ac[2:4], ~select(., -c(created:expired)))
ac[[5]] <- select(ac[[5]], -c(created, modified, expired))
```

Put all five survey's responses on one row per participant

```{r}
ac <- ac %>% reduce(left_join)
glimpse(ac)
glimpse(as_factor(ac))
```

Create a file with codes (hashed ID) and when survey started

```{r eval = FALSE}
select(ac, code, survey_start = created) %>% 
  arrange(survey_start) %>% 
  write_csv(here(str_glue("ID---{Sys.time()}.csv")))
```

Because we only keep telemetry data of players who answered the survey, we'll get a variable with the IDs of ppl who filled survey

```{r}
survey_ids <- pull(ac, code) 
```


### Clean survey data

- Create duration variable
- give some sensible variable names
- assign proper variable types

```{r}
# Duration of survey
ac <- ac %>% 
  mutate(
    survey_duration = ended-created
  )
# Create variables for straightliners by checking if variance within a block of questions is zero
ac$spane_straightliner <- apply(
  select(ac, starts_with("spane_") & !starts_with("spane_acnh")), 
  1, sd, na.rm = TRUE
)
ac$spane_straightliner <- ac$spane_straightliner==0
ac$motivations_straightliner <- apply(
  select(
    ac, 
    starts_with("autonomy_"), 
    starts_with("competence_"), 
    starts_with("related_"), 
    starts_with("enjoymen_"), 
    starts_with("extrinsic_")
  ), 
  1, sd, na.rm = TRUE
)
ac$motivations_straightliner <- ac$motivations_straightliner==0
table(ac$spane_straightliner)
table(ac$motivations_straightliner)

# These are needed as factors
ac <- ac %>% 
  mutate(across(c(sex, played_with_others), as_factor))


# reverse scored items
ac <- ac %>% 
  mutate(
    across(
      c(
        related_not_close,
        enjoyment_attention,
        enjoymen_boring
      ),
      ~ 8 - .x
    )
  )
```

Next, let's create mean indices for the scales.
SPANE has positive affect, negative affect, and an affect balance score (subtract negative from positive).
```{r create-scales}
select(ac, starts_with("spane_"))
    # general spane
ac <- ac %>% 
  mutate(
    spane_positive = rowSums(
      select(
        .,
        spane_positive,
        spane_good,
        spane_pleasant,
        spane_happy,
        spane_joyful,
        spane_contented
      ),
      na.rm = TRUE
    ),
    spane_negative = rowSums(
      select(
        .,
        spane_negative,
        spane_bad,
        spane_unpleasant,
        spane_sad,
        spane_afraid,
        spane_angry
      ),
      na.rm = TRUE
    ),
    spane_balance = spane_positive - spane_negative
  )

  # motivations
ac <- ac %>% mutate(
  autonomy = rowMeans(select(., starts_with("autonomy")), na.rm = TRUE),
  competence = rowMeans(select(., starts_with("competence")), na.rm = TRUE),
  relatedness = rowMeans(select(., starts_with("relatedness")), na.rm = TRUE),
  enjoyment = rowMeans(select(., starts_with("enjoymen")), na.rm = TRUE),
  extrinsic = rowMeans(select(., starts_with("extrinsic")), na.rm = TRUE)
)

  # spane because of playing AC:NH
ac <- ac %>% 
  mutate(
  spane_acnh_positive = rowSums(
    select(
      .,
      spane_acnh_positive,
      spane_acnh_good,
      spane_acnh_pleasant,
      spane_acnh_happy,
      spane_acnh_joyful,
      spane_acnh_contented
    ),
    na.rm = TRUE
  ),
  spane_acnh_negative = rowSums(
    select(
      .,
      spane_acnh_negative,
      spane_acnh_bad,
      spane_acnh_unpleasant,
      spane_acnh_sad,
      spane_acnh_afraid,
      spane_acnh_angry
    ),
    na.rm = TRUE
  ),
  spane_acnh_balance = spane_acnh_positive - spane_acnh_negative
  )
  
  # hours of estimated play
ac <- ac %>% mutate(
  active_play = active_play_hours + (active_play_minutes / 60)
)
```

### Checking

[todo]

## Save file

```{r}
write_rds(ac, here("data/noa/ac.rds"))
```
